---
title: "Exploratory Analysis - Bitcoin price"
author: "Steve White"
date: "10/7/2021"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 2
---

<style type="text/css">
  p,ul,li{
  font-size: 12pt;
}
</style>

```{r setup, include=FALSE}
library(htmltools)
library(tidyverse)
library(magrittr)
library(plotly)
library(tsibble)
library(feasts)
library(fable)

knitr::opts_chunk$set(echo = FALSE, message=FALSE, results = "asis")
plot_res <- readRDS("../data/plot_res.rds")
btc_daily_ts_train <- readRDS("../data/btc_daily_ts_train.rds")

```

## Overview  

This document answers exercise 1 as well as explores the TS analysis framework provided by:
[https://otexts.com/fpp3/arima-r.html](https://otexts.com/fpp3/arima-r.html)

The data analysed is the daily close price of Bitcoin.  On 11, Nov 2021, the first 90% of the periods were taken as a training set.

[Code Available on GitLab](https://gitlabstats-prd/StatsMethods/time-series-network/learning/arima-modelling/lecture-1/-/tree/swhite)  

## Analyse Raw Data {.tabset .tabset-pills}

Are there any unusual observations?  
Years with explosive growth: 2010, 2013, 2017  

```{r}

cat('\n') 
cat("### Observed data  \n")
htmltools::tagList(plot_res$btc_raw)
cat('\n') 
cat("### By year  \n")
htmltools::tagList(plot_res$btc_raw_byyear)
cat('\n') 
# cat("### ACF  \n")
# btc_daily %>% ACF(close) %>% autoplot()
# cat('\n\n') 
# cat("### First Difference  \n")
# p <- btc_daily %>% autoplot(difference(close)) +
#   labs(title="Daily Close Price - 1st Diff",
#        y="USD") 
# htmltools::tagList(ggplotly(p))
# cat('\n\n') 
# cat("### ACF - First Diff  \n")
# btc_daily %>% ACF(difference(close)) %>% autoplot()
# 
# cat('\n\n')

  
```

## {-}

## Box-Cox Transformation {.tabset .tabset-pills}

Should we consider a Box-Cox transformation?
I chose a lambda using Guerro method.

Guerrero, V. M. (1993). Time-series analysis supported by power transformations. Journal of Forecasting, 12(1), 37–48. [DOI]

```{r}

cat('\n') 
cat("### Guerro  \n")
htmltools::tagList(plot_res$btc_guerrero)
cat('\n') 
cat("### Log  \n")
htmltools::tagList(plot_res$btc_log)
cat('\n')


  
```

## {-}

## Differencing {.tabset .tabset-pills}

First difference applied to make stationary.  

```{r, echo=TRUE, results='markup'}


btc_daily_ts_train %>% features(close_guerro, unitroot_ndiffs)
btc_daily_ts_train %>% features(close_guerro, unitroot_kpss)
btc_daily_ts_train %>% features(close_guer_diff, unitroot_kpss)
btc_daily_ts_train %>% features(close_guerro, unitroot_nsdiffs)


```

## {-}

## Plots {.tabset .tabset-pills}


```{r}

cat('\n')
cat("### Guerro Diff  \n")
btc_daily_ts_train %>% autoplot(close_guer_diff)
cat('\n\n')
cat("### ACF PACF  \n")
btc_daily_ts_train %>% ACF(close_guer_diff) %>% autoplot()
btc_daily_ts_train %>% PACF(close_guer_diff) %>% autoplot()
cat('\n\n')
cat("### Lag Plots  \n")
btc_daily_ts_train  %>% gg_lag(close_guer_diff, geom="point")
cat('\n\n')



  
```

## {-}

```{r}
# for(i in snames) {
#   cat(sprintf("## %s  {.tabset .tabset-pills}   \n", i))
# 
#   cat('\n')
#   cat("### Actual-D11-D12  \n")
#   print(htmltools::tagList(hlfs_plots_def$prd[[i]]))
#   cat("\n")
#   cat("### Compare D11  \n")
#   print(htmltools::tagList(d11_combined_plots[[i]]))
#   cat("\n")
#   cat("### Revision Plots  \n")
#   print(htmltools::tagList(revision_plots[[i]]))
#   cat("\n")
#   cat("### Compare D11 Movement  \n")
#   print(htmltools::tagList(d11_combined_plots_mvmt[[i]]))
#   cat("\n")
# 
# 
#   cat("## {-}  \n")
#   
# }

  
```
